\section{Getting started}

\subsection{What \spetc{} is and what it computes}

\spetc{} answers the two questions every observation begins with: \emph{what
signal-to-noise ratio will I reach in a given exposure}, and \emph{how long
must I expose to reach the signal-to-noise ratio I need}. It answers them
for two kinds of measurement: broad-band aperture photometry (measuring the
brightness of an object through a filter) and spectroscopy (dispersing the
light with a slit spectrograph or a slitless grating such as the Star
Analyser and measuring the spectrum). The program is not tied to any
particular observatory: you describe \emph{your} telescope, \emph{your}
camera and \emph{your} sky, and the program predicts what that combination
will deliver on a target of a given brightness and spectral type, at a
given place on the sky, at a given time of a given night.

The prediction is physical, not statistical: the program starts from an
absolutely calibrated template spectrum of a star similar to your target,
scales it to the magnitude you enter, pushes that spectrum through every
element of the light path --- the atmosphere at the current airmass, the
telescope aperture, the filter, the optics, the detector quantum
efficiency --- and counts the photo-electrons that arrive in your aperture
or in each spectral resolution element. It then adds every noise source
that a real detector and a real atmosphere contribute (photon noise from
the object and the sky, dark current, read noise, atmospheric
scintillation, analogue-to-digital quantization) and reports the resulting
signal-to-noise ratio, along with a saturation prediction for the
brightest pixel. The companion document \emph{Physics of the ETC} derives
every formula; this guide tells you which buttons to press and what the
numbers mean.

\subsection{Requirements}

You need Python 3.10 or newer with Tk support. Tk (the graphical toolkit)
is included in the standard installers from \url{python.org} on Windows
and macOS; on Linux it is usually a separate package (for example
\code{sudo apt install python3-tk} on Debian/Ubuntu or
\code{sudo dnf install python3-tkinter} on Fedora). Everything else is
installed with pip. In a terminal, change into the directory where you
unpacked \spetc{} and run:

\begin{verbatim}
python3 -m pip install -r requirements.txt
python3 etc_gui.py
\end{verbatim}

The first command installs the scientific libraries the program is built
on (NumPy, SciPy, pandas, Matplotlib, Astropy and astroquery). The second
starts the program. On Windows, replace \code{python3} with \code{py}.

\subsection{The data directory}

Beside \code{etc\_gui.py} there must be a directory called \code{data}.
It is the program's library and ships ready to use; you never need to edit
it to get started, but you will add to it as you customise the program, so
it is worth knowing what lives inside:

\begin{description}
\item[\code{interpola.db.csv}] The \emph{template catalogue}: one line per
  stellar template, giving its name, spectral type, brightness
  calibration and the file that holds its spectrum. The template list you
  see in the GUI is this file, read at start-up. Its exact format is
  specified in Sect.~\ref{sec:file-catalog}.
\item[\code{bpgs/}] The template spectra themselves: absolutely calibrated
  stellar spectra from the Bohlin CALSPEC/BPGS libraries, one file per
  star, covering the main spectral types from O to M plus the Sun and
  Vega. You can add your own (Sect.~\ref{sec:file-templates}).
\item[\code{filters.list} and \code{filters/}] The list of available
  filters and their transmission profiles in the Spanish Virtual
  Observatory (SVO) XML format --- Johnson--Cousins, Sloan, 2MASS and
  others, each carrying its own photometric calibration. The special
  \code{BLANK.xml} profile represents ``no filter at all''
  (Sect.~\ref{sec:file-filters}).
\item[\code{qe.dat}] The default detector quantum-efficiency curve: how
  many electrons the sensor produces per incoming photon, as a function of
  wavelength (Sect.~\ref{sec:file-curves}).
\item[\code{earth\_atmospheric\_transmission.fits}] Optional: a measured
  atmospheric transmission table (the shipped one is the Paranal
  extinction product). When absent, a built-in broad-band extinction
  table is used instead.
\end{description}

Two more files live beside the program itself rather than in \code{data}:
\code{etc\_sites.json}, your personal list of observing sites, and
\code{etc\_user\_config.json}, the saved state of every control in the
window. Both are created and updated automatically.

\subsection{First launch and configuration persistence}

On the first start you get the default configuration: the Asiago (Cima
Ekar) observatory, a 358-mm telescope, and Vega as the target template.
Every control in the window --- every number, every selector --- is saved
to \code{etc\_user\_config.json} after each successful calculation and
when you quit, and restored on the next start. There is no ``save
settings'' button because saving is automatic; to transfer your setup to
another computer, copy that file (and your instrument profile, see
Sect.~\ref{sec:profiles}) along with the program.

\subsection{A first calculation, explained}

The fastest way to understand the program is to run it once. The
following walk-through predicts the photometry of a 10th-magnitude star
tonight.

First, tell the program \emph{where} you are. In the leftmost column,
open the site selector and choose one of the built-in observatories, or
type your latitude (degrees, north positive), longitude (degrees, east
positive) and elevation (metres) directly into the three fields. The
site matters more than it may seem: the latitude and longitude determine
when and how high your target rises, the elevation sets the atmospheric
pressure used for refraction and the air mass of turbulence responsible
for scintillation noise.

Second, tell it \emph{when} you observe. Set the date and the UT time.
If you think in local time, note the timezone selector: choose an IANA
timezone name such as \code{Europe/Rome} and the program handles daylight
saving for you in all displayed local times; the calculation itself
always runs on UT.

Third, tell it \emph{what} you observe. Type the target coordinates into
the RA and Dec fields. Both decimal degrees (\code{279.2347}) and
sexagesimal (\code{18:36:56.3}) forms are accepted. If you know only the
object's name, type it into the name field and press \textsf{Resolve
name}: the program queries the SIMBAD database (this is the only feature
that uses the internet) and fills the coordinate fields for you. Then
enter the target's magnitude --- say 10.0 --- and check that the
\emph{reference filter} selector says in which band that magnitude is
expressed (V, by default) and in which system (Vega, by default).

Fourth, choose a template. In the template list, pick a star of roughly
the same spectral type as your target --- for a Sun-like target choose a
G2V, for a hot star a B or A star. The template provides the \emph{shape}
of the spectrum; your entered magnitude provides the level. The panel
shows a preview of the selected spectrum so you can see what you picked.

Finally, press the green \textsf{Run ETC} button. Three things appear:
the \textbf{Results window} with the numbers (Sect.~\ref{sec:results}),
the \textbf{plot window} with S/N versus time across the night, the
altitude track and the sky path of the target, Sun and Moon, and the
status line reporting completion. The red marker in the plots is the
time you selected; the curve around it shows how the same observation
would fare earlier or later in the night, with the sky brightness, the
airmass and the Moon recomputed at every time step.

If the button is grey rather than green, the program is telling you that
the current combination cannot be computed --- most often because the
selected template does not overlap the selected filter. The message above
the response preview says exactly what is wrong; fix the selection and
the button turns green.
