\section{The main window}

The window is organised in five columns, left to right.

\subsection{Observation, site and target}

Site selector (built-in observatories merged with your persistent
\code{etc\_sites.json}; \textsf{Save current site} and \textsf{Import site
file} manage it), geographic coordinates and elevation, date and UT time of
the planned observation, the observing-interval selector, the time zone
(IANA name with DST, e.g.\ \code{Europe/Rome}, or a fixed UTC offset), and
the target: RA/Dec fields (decimal degrees or \code{hh:mm:ss} /
\code{dd:mm:ss}) with the optional SIMBAD \textsf{Resolve name} button.
Running the ETC itself never uses the network: only the coordinates in the
fields count.

\subsection{Telescope, detector and conditions}

Telescope diameter, central obstruction, focal length and scalar optics
throughput; optional calibrated instrument-response and slit
resolving-power curves (two-column files with an explicit wavelength unit);
detector pixel size, gain, full well, ADC bit depth, read noise and dark
current; seeing FWHM; the sky-background mode (\code{ing} = physical
night/twilight/day model, \code{fixed\_ab} = your own observed sky surface
brightness); and the \textbf{PSF model} (\code{gaussian} or \code{moffat}
with its $\beta$; Moffat reproduces the wings of real star images and is
recommended). \textsf{Save/Load instrument profile} stores the whole column
as a portable JSON with its QE and response curves.

\subsection{Calculation settings}

Photometry or spectroscopy; exposure time; optional \textbf{target S/N}
(when set, the ETC returns the required exposure instead, flagging
saturation). The photometry panel adds the aperture radius and the
\textbf{source geometry}: \code{point}, or \code{extended} with the source
area in arcsec$^2$ for galaxies, nebulae and planets (the magnitude you
enter stays the \emph{total} integrated magnitude). The spectroscopy panel
holds the resolving power, slit width, extraction height, sampling, the
slitless parameters (Sect.~\ref{sec:spec-guide}), the wavelength range and
the S/N reference wavelength, and the \textbf{slit orientation}
(\code{parallactic} or \code{fixed}).

\subsection{Filter and template selection}

Two independent filters: the \textbf{reference} filter, in which your
magnitude is expressed (Vega or AB), and the \textbf{observing} filter,
physically in the beam for both photometry and spectroscopy (spectroscopy
defaults to \code{BLANK}, the unfiltered response). The panel previews the
observing transmission curve with its pivot wavelength, width and zero
point; the template selector searches the catalogue by name and previews
the selected flux distribution; \textsf{Validate V and B$-$V} checks the
template's synthetic colours against the catalogue.

\subsection{Run actions}

\textsf{Run ETC}, the live observatory status (Sun/Moon altitude, current
airmass of the target), and \textsf{Exit} (which saves the configuration).
