\section{Spectroscopy step by step}
\label{sec:spec-guide}

\subsection{Slit spectroscopy}

\textbf{Step 1 — establish the resolving power.} Everything in a
spectroscopic prediction hangs on the resolving power
$R = \lambda/\Delta\lambda$, because it fixes how the light is divided
among resolution elements. You have three ways to supply it, in
increasing order of fidelity. If you know $R$ from your spectrograph's
documentation, type it into \textsf{Slit resolving power R} and you are
done. If you know your \emph{hardware} instead --- most amateurs do ---
fill the three geometry fields (grating grooves per mm, collimator focal
length, camera focal length) and press \textsf{Compute R from geometry}:
the program combines them with the telescope focal length, your slit
width and the seeing through the grating equation, fills the R field,
and tells you whether the resolution is limited by the slit or by the
seeing --- if it says ``seeing-limited'', a narrower slit would only lose
light without sharpening the spectrum. Finally, if you have a
\emph{measured} $R$-versus-slit-width curve for your instrument, load it
in the telescope column and it silently overrides the nominal value
within its calibrated range.

\textbf{Step 2 — describe the extraction.} The slit width (arcsec)
controls how much starlight enters; the extraction height (arcsec) is
the length along the slit that your reduction software will sum over.
Both cost and gain: a wide slit and tall extraction capture more of the
seeing-blurred star but admit proportionally more sky, and the ETC
accounts for both directions exactly, using the PSF model you selected.
\textsf{Pixels per resolution element} is your detector sampling
(typically 2--3); it fixes how many pixels' worth of read noise each
spectral bin carries.

\textbf{Step 3 — decide the slit orientation.} Atmospheric refraction
disperses every star into a tiny vertical spectrum: at airmass 1.5 the
blue image at 4000\,\AA{} sits about $1''$ from the red image at
6600\,\AA{}. If your slit is aligned with that vertical direction ---
the \emph{parallactic angle}, reported for every time sample in the
Results window --- all images fall along the slit and nothing is lost.
If the slit sits at any other angle, the blue and red images drift out
of it sideways. Choose \code{parallactic} if you rotate your slit during
the night (or observe near the meridian, where the effect vanishes);
choose \code{fixed} if your slit angle is fixed, and the program will
charge the worst-case wavelength-dependent loss --- tens of percent at
the blue end at airmass 1.5--2, exactly the classical Filippenko effect.
Comparing a run in each mode shows you precisely what slit rotation
would buy.

\textbf{Step 4 — set the wavelength range and reference.} The range
(min/max, \AA{}) is the span the spectrum table will cover; the
\textsf{S/N reference wavelength} is the single wavelength at which the
time-series curve and the stack plan are evaluated --- put it where your
science is (H$\alpha$ at 6563 for emission-line work, 5500 for general
use).

\textbf{Step 5 — run and read.} The plot window adds two spectral
panels: the detected source rate and the S/N, both \emph{per resolution
element}, with a manual time slider that recomputes the whole spectrum
at other times of the night. The Results table gives the full
wavelength-by-wavelength budget (Sect.~\ref{sec:results}).

\subsection{Slitless spectroscopy and the Star Analyser}

In slitless spectroscopy there is no slit: every star in the field is
dispersed directly, resolution is set by the seeing rather than by a slit,
and no slit losses occur --- which is why a Star Analyser on a modest
telescope reaches surprisingly bright-looking spectra, at the price of
low resolution and of sky light entering at every wavelength.

Mode \code{slitless} offers two ways to describe the disperser. For a
\textbf{Star Analyser 100 or 200} (or any transmission grating in the
converging beam), enter the groove density (100 or 200 lines/mm) and the
distance from the grating to the sensor in millimetres --- for a filter-%
wheel mounting this is roughly the wheel-to-sensor distance; measure it
once. From these two numbers and your pixel size the program computes
the dispersion itself: an SA100 at 42\,mm on 4.63\,\si{\micro\metre}
pixels gives 11.0\,\AA/pixel, matching the published calibration. The
optional grating efficiency (a realistic first-order value is
$\sim$0.5) scales source and sky alike. Alternatively, leave the groove
density at 0 and type a known dispersion in \AA/pixel directly.

The resolution element is computed from the seeing, the intrinsic
grating blur and the atmospheric-dispersion smear combined; with
$2$--$3''$ seeing expect $R \sim 100$--200, and note that longer
exposures do \emph{not} sharpen it --- only better seeing or a longer
grating distance do. The cross-dispersion extraction height plays the
role the extraction height plays in slit mode. Slitless results are
flagged \emph{experimental} until validated against a calibrated
instrument; treat absolute count levels with one raised eyebrow and
relative comparisons with confidence.

\subsection{Getting a science answer: S/N, $\sigma$(EW) and the stack plan}

All spectroscopic counts and S/N are \textbf{per resolution element},
not per pixel --- when comparing with other tools, check which
convention they use, since the difference is a factor $\sqrt{N_{\rm
pix}}$. Each result row also carries $\sigma$(EW), the equivalent-width
uncertainty at that wavelength from the Cayrel relation, in
milli-\AA{}ngstr\"om. This converts S/N into the language of actual
measurements: a line is securely measurable when its expected
equivalent width exceeds roughly $3\,\sigma$(EW). Concretely, with
$\sigma$(EW) = 15\,m\AA{} at H$\alpha$, the 50--500\,m\AA{} emission
variations of a Be star are easy science, a 100\,m\AA{} interstellar
line is solid, and a 20\,m\AA{} photospheric line is out of reach ---
one number, read directly from the table, settles what your night can
and cannot do. The value at the reference wavelength is repeated in the
label under PHOTOMETRY and in the outputs tab, and the stack planner
works exactly as in photometry, applied to the reference resolution
element.
