\section{On-disk formats}

\subsection{\code{data/} directory}

\begin{description}
  \item[\code{interpola.db.csv}] Header line then comma-separated rows:
    \code{nrow, Vmag, B-V, name, spt, filename}; \code{filename} is
    relative to \code{data/} (subdirectories allowed, e.g.\
    \code{bpgs/bpgs\_92.ascii}). \code{Vmag} is the CALSPEC visual
    magnitude represented by the file ($m_{v0}$).
  \item[template spectra] Two-column ASCII (wavelength [\AA{}], $\flam$)
    or CALSPEC/STScI-style FITS binary tables
    (\code{WAVELENGTH}/\code{FLUX} and common variants; \code{TUNIT} in
    Angstrom, nm or micron; extra columns ignored), dispatched by
    extension. Non-finite and non-positive rows are dropped; wavelengths
    are sorted and deduplicated on load. \code{add\_template.py} imports
    files or directories, computing the synthetic $m_{v0}$/$B-V$ and
    appending the catalogue rows.
  \item[\code{qe.dat}] Two-column wavelength/QE; the wavelength unit is
    declared in the configuration (\code{Angstrom}/\code{nm}/\code{um}).
  \item[\code{filters.list}] One entry per line naming SVO profiles;
    entries with or without the \code{filters/} prefix and \code{.xml}
    suffix all resolve. Each logical filter needs \code{<name>\_Vega} and
    \code{<name>\_AB} VOTables carrying \code{MagSys}, \code{ZeroPoint}
    [Jy], \code{DetectorType}, \code{WavelengthUnit} and the transmission
    table. \code{BLANK.xml} (unity 3000--26000 \AA{}, AB) is optional and
    auto-detected.
  \item[\code{earth\_atmospheric\_transmission.fits}] Optional; accepted
    layouts in Sect.~2 (\code{spectral\_utils}).
  \item[CMOS gain tables] One row per gain setting, whitespace- or
    comma-separated, \code{\#} comments allowed:
    \code{setting  e-/ADU  read\_noise\_e  full\_well\_e}. Loaded through
    \code{detector.load\_gain\_table}; the GUI persists the path in
    \code{gain\_table\_path} and restores it at start-up.
  \item[Batch run descriptions] JSON, see \code{examples/}: top-level
    \code{mode}, \code{telescope}, \code{detector}, \code{atmosphere},
    \code{sky} (\code{fixed\_ab}/\code{sqm}), \code{target},
    \code{photometry} or \code{spectroscopy} blocks, optional
    \code{target\_snr} and \code{stack\_sub\_exposure\_s}.
\end{description}

\subsection{Configuration and profiles}

\code{etc\_user\_config.json} is a flat string-to-string map of every
registered GUI variable plus \code{schema\_version}; it is rewritten after
each successful run and on exit. Instrument profiles are the same keys
restricted to the instrument subset (\code{\_instrument\_profile\_keys}),
saved with relative copies of the QE, response and slit-$R$ curves beside
them. \code{etc\_sites.json} holds user sites
(\code{name/lat/lon/elev/utc\_offset\_h/timezone/sky\_ab\_mag\_arcsec2});
\code{observatories.json} the built-in presets. When adding a new
persistent setting, register it with \code{self.\_var("key", default)} ---
persistence is automatic; bump \code{schema\_version} only for
incompatible changes.
